\documentclass{amsart}
\usepackage{parskip}
\usepackage{ytableau}
\DeclareMathOperator{\Ind}{Ind}

\usepackage{stmaryrd}
\usepackage[inline]{showlabels}
%Title Style
\usepackage{mathtools}
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@settitle}{\uppercasenonmath\@title}{}{}{}
\patchcmd{\@setauthors}{\MakeUppercase}{}{}{}
\patchcmd{\section}{\scshape}{}{}{}
\makeatother
\makeatletter
\patchcmd{\@maketitle}
  {\ifx\@empty\@dedicatory}
  {\ifx\@empty\@date \else {\vskip2ex %vertical position of date
  \centering\footnotesize\@date\par\vskip1ex}\fi
   \ifx\@empty\@dedicatory}
  {}{}
\patchcmd{\@adminfootnotes}
  {\ifx\@empty\@date\else \@footnotetext{\@setdate}\fi}
  {}{}{}
\makeatother


%Page Style
\usepackage{mathabx}
\usepackage[colorlinks,backref=page,citecolor=blue]{hyperref}
\usepackage{accents}
\renewcommand{\familydefault}{ppl}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\marginparwidth}{0in}
\setlength{\marginparsep}{0in}
\setlength{\marginparpush}{0in}
\setlength{\topmargin}{0.3in}
\setlength{\headsep}{14pt}
\setlength{\footskip}{.3in}
\setlength{\textheight}{8.0in}
\setlength{\textwidth}{5.8in}
\setlength{\parskip}{4pt}
\linespread{1.2}



\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{restate}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{conj}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{example}[theorem]{Example}
\newtheorem{definition}[theorem]{Definition}
\DeclareMathOperator{\YY}{\mathbb{Y}}
\DeclareMathOperator{\Y}{\bf{Y}}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sh}{shape}
\DeclareMathOperator{\SYT}{SYT}


\title{Math 206A Problem Sets}

\begin{document}
	\maketitle
\begin{itemize}
\item Submit a genuine attempt for each assigned problem set. Perfect solutions are not required for full credit, but your work must be clearly typesetted.
\item Printed submissions are preferred. Email submissions are allowed in special circumstances.
\item Collaboration is encouraged. However, write up your solutions independently and acknowledge any collaborators.
\item \textbf{AI policy:} You may use LLMs to find references or write codes. Do not use AI to directly solve or write solutions for this problem set.
\item  {\color{red}[Add.] Here's a more rigorous definition of a genuine attempt  $:\geq$ perfect solutions for half of each problem set. } 
\item {\color{red}[Add.] Numbers after each problem shows difficulty, $3+>3>3->2+>\cdots$. Grades don't depend on difficulty.} 
\end{itemize}
	
	
	
	
	
		
		\section{Homework Assignment 1 (Due Oct 10 F)}
	
	\begin{problem}[2+]
		Recall that $\mathbb{Y} = \mathbb{Q}\text{-span}(\bf{Y})$, where $\bf{Y}$ is the Young's lattice, and the up/down operators $U,D$ defined by $U(\lambda)=\sum_{\lambda\lessdot \mu}\mu$ and  $D(\lambda)=\sum_{\mu\lessdot \lambda}\mu$. Prove that 
		\[[D,U]=\text{id}\]
	\end{problem}
	
	\begin{problem}[2-]
		Recall the Weyl algebra $\mathcal{W}$ is the $\mathbb{Z}$-algebra with unit $1$ and generators $U,D$ with relations $[D,U]=1$. Rewrite $D^nU^n$ as an $\mathcal{W}$-element such that no $D$ appears before an $U$. What is the identity coefficient?
		
		For example, $D^2U^2=D(UD+1)U=DUDU+DU=(UD+1)(UD+1)+(UD+1)=UDUD+3UD+2=U(UD+1)D+3UD+2=U^2D^2+4UD+2$. The identity coefficient is $2$.
	\end{problem}
	
	\begin{problem}[2]
		Prove that $\lambda\trianglelefteq \mu$ if and only if $\mu'\trianglelefteq \lambda'$. Here $\lambda'$ is the conjugate of $\lambda$.
	\end{problem}
	
	\begin{problem}[2]
		Prove that
		\[\sum_{k\geq 0} h_k t^k = \prod_{i\geq0}{1\over 1-x_it}\]
	\end{problem}
	
	\begin{problem}[2+]
		Is the power series $f=\prod_{i\geq 1}(1+x_i+x_i^2)$ symmetric? If so, expand in the $e$-basis. [Hint: You can use (without proof) the change of basis matrix between $m_\lambda$ and $e_\lambda$.]
	\end{problem}
		\begin{problem}[2]
		Expand $h_3e_4$ in the Schur basis. 	
	\end{problem}
	\begin{problem}[2+]
		[Bonus] Expand $h_me_n$ in the Schur basis. You may start experimenting with {\tt{SageMath}} to make a conjecture.
	\end{problem}
	
	
	\section{Homework Assignment 2 (Due Oct 20 M)}
	\begin{problem}[2+] 		Prove that if $\lambda,\mu\vdash n$ and $\lambda\trianglelefteq \mu$, then $K_{\lambda\mu}\neq 0$.
	\end{problem}
\begin{problem}[2]
	Prove that $s_\lambda\cdot s_{\square}=\sum_{\lambda \lessdot \mu}{s_\mu}$. Here $s_{\square}=m_1=x_1+x_2+\cdots$, and $\lessdot$ denotes the covering relation in Young's lattice.
\end{problem}

{\color{blue}Rmk: Can you define a $\mathcal{W}$-action on $\Lambda_{\mathbb{Q}}$? By P 2.2, we get an action of $U$ on $\Lambda$ via multiplication by $s_{\square}$. What should be the action of $D$ on $\Lambda_{\mathbb{Q}}$ such that the map $ \lambda \mapsto s_\lambda$ is an isomorphism from $\mathbb{Y}$ to $\Lambda_{\mathbb{Q}}$ as $\mathcal{W}$-modules?}

\begin{problem}[2+]
	Let $u_i:\YY\to \YY$ be the operator of adding a box to the $i$-th row when possible. Define 
	\[H_k=h_k(u_1,u_2,\cdots)=\sum_{1\leq i_1\leq i_2\leq\cdots\leq i_k}u_{i_1}u_{i_2}\cdots u_{i_k}\]
	The operators $u_i$ satisfy the relations 
	\begin{align*}
	u_iu_j&=u_ju_i\text{\quad if}|i-j|\geq 2\\
	u_iu_{i+1}u_i&=u_{i+1}u_iu_i\\
	u_{i+1}u_{i+1}u_i&=u_{i+1}u_iu_{i+1}\\
	u_{i+1}u_{i+2}u_{i+1}u_i&=u_{i+1}u_{i+2}u_iu_{i+1}	
	\end{align*}
	Classify the relations of $\{H_i|i\in\mathbb{N}\}$.
	
	(Bonus) What about 
	$$E_k=e_k(u_1,u_2,\cdots)=\sum_{1\leq i_1<i_2<\cdots<i_k}u_{i_1}u_{i_2}\cdots u_{i_k}$$
	
	
[Hint: How does $H_k$ act on $\YY$? You may or may not need to use the relations of the $u_i$'s.] 
\end{problem}
\begin{problem}[3+]
	
	Define $H'_k=h_k(d_1,d_2,\cdots)$. What relations do $\{H_i,H'_j:i,j\in\mathbb{N}\}$ satisfy?
\end{problem}
\begin{problem}[3]
	Let $P(w)$ denote the row-insertion tableau, and $P'(w)$ denote the column-insertion tableau. Prove that $P(w)=P'(w_0w)$, where $w_0w=(w_n,w_{n-1},\cdots,w_2,w_1)$.
\end{problem}


	
	\section{Homework Assignment 3 (Due Oct 29 W)}
	\begin{problem}[2]
	Define $A(x)=\cdots (1+x u_3)(1+x u_2)(1+x u_1)$ and $B(x)= (1+x d_1)(1+x d_2)(1+x d_3)\cdots$. Prove that $B(y)A(x)=A(x)B(y)(1-xy)^{-1}$. 
	[Hint: first show that $(1+a)(1-ba)^{-1}(1+b)=(1+b)(1-ab)^{-1}(1+a)$ for any non-commutative $a,b$.]
\end{problem}
\begin{problem}[2+]
	Let $P$ be a finite poset, and $a_k(P)$ the size of maximal $k$-antichain of $P$. Prove that
	\[a_k(P)=a_1(P\times[k])\]
	where $[k]$ is the chain of $k$ elements.
\end{problem}
\begin{problem}[Bonus, 3+]
	Let $P$ be a finite poset. Let $C=(C_1,C_2,\cdots,C_l)$ be a partition of $P$ into chains. 
	
	(i) Show that for any $C$, we have
	\[a_k\leq \sum_i \min(|C_i|,k)\]
	
	A chain-partition $C$ is called \emph{$k$-saturated} if the equality  holds. 
	
	(ii) Prove that for any $P$ there exists a $k$-saturated chain partition for all $k$. 
	
	[Hints: (1) The case $k=1$ is the Dilworth theorem, which you can use without proving (2) use Problem 3.3]
\end{problem}
\begin{problem}[2+]
		Denote $\SYT_{\lambda\setminus \mu}$ the set of all skew standard Young tableaux of shape $\lambda\setminus \mu$. Fix a $\mu\in\Y$ and define	\(D_\mu=\{(T_1,T_2)\in \SYT_{\lambda\setminus \mu}:|\lambda\setminus \mu|=n\}\).
		What is $\# D_\mu$? Note that the RS bijection implies that $\# D_{\emptyset}=n!$.
		
		{\color{red}correction: You only need to find a lower bound for $\# D_\lambda$.}
	
	[Hint: use growth diagram.]
	\end{problem}

\begin{problem}[Bonus, 5]
	Recall that a permutation is an involution, i.e. $w=w^{-1}$, if and only if $P(w)=Q(w)$. The Bruhat order on $S_n$ induces a partial order on involutions, thus a partial order on SYT's. Give a description of this partial order on SYT's.
\end{problem}
	\section{Homework Assignment 4 (Due Nov 7 F)}

\begin{problem}[2]Consider
\[
A = \begin{pmatrix}
-1 & 1 & 2 \\
5 & 0 & 2 \\
-3 & 2 & 1
\end{pmatrix},
\qquad
B = \begin{pmatrix}
-4 & 4 & 4 \\
5 & -3 & -4 \\
5 & 5 & -4
\end{pmatrix}.
\]
And let $E_A=\langle A_1\rangle \subset \langle A_1,A_2\rangle\subset \langle A_1,A_2,A_3\rangle ^3 $ and let $E_B=\langle B_1\rangle \subset \langle B_1,B_2\rangle\subset \langle B_1,B_2,B_3\rangle ^3 $. Compute the relative position of $E_A,E_B$.

\end{problem}

\begin{problem}[2+]
	(i) Compute $\mathfrak{S}_{321654}$.
	
	(ii) Which Schubert polynomials are monomials?
\end{problem}
\begin{problem}[3]
	Let $\text{GP}_{k,n}$ denote the set of all $k$-Grassmanian permutations in $S_n$. For $w\in \text{GP}_{k,n}$, denote $\lambda(w)$ the partition associated to it. Prove that $\mathfrak{S}_w=s_{\lambda(w)}(x_1,\cdots,x_k)$.\end{problem}
\begin{problem}[Bonus, 4]
	Let $\tau:S_n\to S_{n+1}$ be the map that sends $w$ to $u$  where $u_1=1$ and $u_{i}=w_{i-1}+1$. For any $w\in S_n$, define $F_w=\lim_{k\to\infty}\mathfrak{S}_{\tau^k(w)}$. Prove that $F_w$ is symmetric
\end{problem}
	\section{Homework Assignment 5 (Due Nov 17 M)}
	\begin{problem}[5+]
	Read Chapter 1 of Sagan's book
\end{problem}
\begin{problem}[1]
	Let $w\in S_n$ and $w'\in S_{n+1}$ such that $w'(i)=w(i)$ for $i\in[n]$ and $w'(n+1)=n+1$. Show that $\mathfrak{S}_w=\mathfrak{S}_{w'}$.
\end{problem}
\begin{problem}[1]
	Use Growth diagram to compute the inverse RSK of $(T_1,T_2)$ where $T_1=\ytableaushort{1134,22,3} $
	and $T_2=\ytableaushort{1223,23,3}$
\end{problem}
\begin{problem}[2]
	Can you construct a permutation representation for the dihedral group $D_{2n}=\{s^2=r^n=1,srs=r^{-1}\}$. Is it faithful?
\end{problem}
\begin{problem}[Bonus, 2]
Recall the action of the (right) divided difference operator on the vector space $\text{span}\langle S_n\rangle $:
\[\partial_i(w)=\begin{cases}
	ws_i & \text{ if }ws_i<w\\
	0& \text{ otherwise}
\end{cases}\]
We also define the left divided difference operator $\sigma_i$:
 \[\sigma_i(w)=\begin{cases}
	s_i w & \text{ if }s_i w<w\\
	0& \text{ otherwise}
\end{cases}\]
Note that both of these operators satisfy the Nil-Coxeter relations.

Can you describe the action of these operators on (the vector space of) ${\bf Y}_{k \times (n-k)}\cong GP_{k,n}\subset S_n$. Here ${\bf Y}_{k \times (n-k)}$ is the set of all partitions that fit inside a $k\times (n-k)$ rectangle.
\end{problem}



	\section{Homework Assignment 6 and 7 (Due Dec 5 F)}
\begin{problem}
	Find all representations of $D_{2n}$ up to equivalence. (Hint: $C_n$ is a normal subgroup of $D_{2n}$).
\end{problem}
\begin{problem}
	For any $w\in S_n$, prove that $\chi^{(n-1,1)}(w)=\#\{\text{fixed points of }\pi\}-1$
\end{problem}
	\begin{problem}
		Suppose we have groups $K\leq H\leq G$ and $\rho:K\to\text{GL}(\mathbb{F}^n)$ a representation of $K$. Prove that
		\[\Ind_{K}^G\rho=\Ind_H^G \Ind_K^H\rho\]
	\end{problem}
\begin{problem}Compute the Specht module for $\lambda=(n-1,1)$. What's its dimension?
\end{problem}

\begin{problem}
	
\end{problem}

\begin{problem}
	
\end{problem}



	
	
	
	
	
	
	
	
\end{document}