\documentclass{amsart}
\input{command.tex}
\usepackage[margin = 1in]{geometry}
\begin{document}
	\leftline{\bf MATH 4242 Quiz 2}\rightline{Name:\underline{\quad\quad\quad\quad\quad\quad\quad\quad}}
	\rightline{Student Id:\underline{\quad\quad\quad\quad\quad\quad\quad\quad}}
	
\vspace{2em}


\begin{enumerate}
\item Prove or provide a counterexample. Does $\{(1,1),(2,5)\}$ form a basis of $\mathbb{R}^2$? (3 pts)
\begin{proof}
	Yes. First we show that $(1,1)$ and $(2,5)$ is linearly independent, i.e. the only way to write $(0,0)$ as a linearly combination of them is using $0$ coefficient. This is equivalent to solving the system of equations.
	\[x+y =0\]
	\[2x+5y=0\]
	Using Gaussian elimination, we found only one solution which is $x=y=0$. Therefore they are linearly independent.
	
	We know that $\dim \rr^2 =2$, and linearly independent list of vectors of the correct size for a basis, so $(1,1),(2,5)$ form a basis of $\rr^2$.
\end{proof}
\item Are the vectors $(1,3),(0,-1),(1,1)\in\rr^2$ linearly independent? Explain your answer. (3 pts)
\begin{proof}
	No. $\rr^2$ has dimension two, so every basis of $\rr^2$ has size two. But we know that if the three vectors were linearly independent, then they extend to a basis of $\rr^2$ of at least three vectors\footnote{Using the lemma that linearly independent list of vectors extends to a basis.}, which is a contradiction.
\end{proof}
\item What's the dimension of $V = \{(x,y,z)|x+y+z = 0\}$, as a subspace of $\rr^3$? (4 pts)
\begin{proof}
	$V$ an be re-written as
	\[V=\{(x,y,-x-y)|x,y\in\rr\}\]
	Every vector in $V$ of the form $(a,b,-a-b)$ can be uniquely written as $a(1,0,-1)+b(0,1,-1)$. Therefore $V = \s\{(1,0,-1),(0,1,-1)\}$. These two vectors are linearly independent, so they form a basis of $V$. Thus $\dim V = 2$.
\end{proof}
\end{enumerate}

%\noindent Part a (4 pts)

%Prove that both $U$ and $W$ are subspaces of $V$.
%\vspace{6cm}
%
%\noindent Part 2 (6 pts)
%
%Describe their sum $U+W$. Is it a direct sum?

\end{document}