If $w_0$ is the largest element in a parabolic subgroup $S_{[i,j]}$, e.g. $143256$ in $[2,4]$.

\begin{prop}
	$C'_{w_0}=\sum_{x\leq w_0}T_x$
\end{prop}
\begin{proof}
	$P_{x,w} = P_{sx,w}$ if $s$ is left descent of $w$ but not of $x$.
\end{proof}

What if instead of factoring a product of $C'$ into $C'_s$'s, factor into $C'_{w_0}$'s where $w_0$ is longest in a certain parabolic subgroup.

Denote $C'_{[i,j]}=\sum_{x\text{ is a permutation on }[i,j]}T_x$.
We have 
\[C'_{s_1s_2} = C'_{s_1}C'_{s_2}\iff C'_{s_1s_2}=C'_{[1,2]}C'_{[2,3]}\]
\begin{example}
	\[C'_{[1,2]}C'_{[2,4]}=(T_{s_1}+1)(T_{s_2s_3s_2}+T_{s_2s_3}+T_{s_3s_2}+T_{s_3}+T_{s_2}+1)=C'_{4132}\]
\end{example}
\begin{example}
	Consider $C'_{[1,3]}C'_{[2,4]}$. From the picture, the longest permutation we can get is $4312$.
	$$(1+q)C'_{4312}=C'_{[1,3]}C'_{[2,4]}$$
\end{example}
Avoiding $3412$ and $4231$ means that the Schubert variety is smooth, and thus Kazhdan-Lusztig polynomial equals to $1$, in which case we can obtain a factorization formula for $C'$ easily. Let's look at when we don't avoid them.
\begin{example} By sorting, the candidate for $3412$ is
	 $C'_{3412}=C'_{[2,3]}C'_{[1,2]}C'_{[3,2]}C'_{[2,3]}$, which is true by Billey-Warrington.
	 
	 For 4231, there are two candidates $C'_{[1,2]}C'_{[2,4]}C'_{[1,2]}$ or $C'_{[3,4]}C'_{[1,3]}C'_{[3,4]}$. It turns out that both are correct.
	 
	 The product on the right hand side is
	 \[(T_{s_1}+1)(T_{s_2s_3s_2}+\cdots+1)(T_{s_1}+1).\]
	 Only need to look at the ``non-reduced part'', which gives $(q+1)(T_{s_1+1})+(q+1)(T_{s_1}+1)T_{s_3}$. Can be checked by length analysis.
\end{example}
\begin{example}
	In $S_5$, $C'_{45312}$ doesn't work.
\end{example}
\begin{theorem}[Agrawal-Sotirov]
	Product of $C'_{[i,j]}$-s is equal to sum of flows (through corresponding heaps) $q^{\#\text{defects}}$
\end{theorem}

\begin{theorem}
	If $k_1,\cdots,k_l$ are ``overlap'' sizes in a heap. We can factor $\prod[k_i]_q!$. If degree of $q$ in the result satisfies $\leq {1\over 2}(\ell(w)-\ell(x)-1)$ condition, the result is $C'_w$.
\end{theorem}