% Set document class and font size
\documentclass[letterpaper, 11pt,times]{article}

\usepackage{calc}
%%
\usepackage{enumerate}
\newcommand\reverselabel[1]{%
  \def\theenumi{}%
  \renewcommand\makelabel{\makebox[\dimexpr\labelwidth-3pt\relax][r]{%
    \the\numexpr#1-\value{enumi}+1\relax}}}
    

%%
\usepackage[utf8]{inputenc}
%\usepackage{CJKutf8}
\newcommand{\mycolor}[0]{\color{RoyalBlue}}

% Package imports
\usepackage{setspace, longtable, graphicx, hyphenat, hyperref, fancyhdr, ifthen, everypage, enumitem, amsmath, setspace, comment}

% --- Page layout settings ---

% Set page margins
\usepackage[left=1in, right=1in, bottom=1in, top=1in]{geometry}

% Set line spacing
\renewcommand{\baselinestretch}{1.15}



% --- Page formatting --- \color{OliveGreen}

% Set link colors
\usepackage[dvipsnames]{xcolor}
\hypersetup{colorlinks=true, linkcolor=OliveGreen, urlcolor=OliveGreen}

% Set font to Libertine, including math support
\usepackage{libertine}
\usepackage[libertine]{newtxmath}

% Remove page numbering
\pagenumbering{gobble}




% --- Document starts here ---


\newlength{\myl}
\settowidth{\myl}{\quad }
\newcommand{\ind}[0]{\hspace{\myl}}
\newlength{\newl}

\newcommand{\paper}[2]{
\settowidth{\newl}{\textbf{#1. \ }}
\noindent
\hspace{\dimexpr\myl - \newl}
\textbf{#1. #2}
}

\begin{document}

% Name and date of last update to this document
\noindent{\hspace{0.3cm}{\Huge{Sylvester W. Zhang}}
\hfill{\it\footnotesize Updated \today}\hspace{0.5cm}}\vspace{0.2cm}\\
%\begin{CJK}{UTF8}{gkai}{{\color{white}z }\Huge张文泽 }\end{CJK}




% --- Contact information and other items ---

\vspace{0.5cm} 
% --- Start the two-column table storing the main content ---

% Set spacing between columns

\setlength{\tabcolsep}{8pt}

% Set the width of each column
\begin{longtable}{p{1in}p{5in}}
% --- Personal info

\nohyphens{\mycolor{Personal Info}}& \textbf{Email}: \href{mailto:sylvesterzhang@math.ucla.edu}{sylvesterzhang@math.ucla.edu} \\&\textbf{Homepage}: \href{http://www.sylvesterzhang.com}{http://www.sylvesterzhang.com}
\\
&\textbf{Address:} 520 Portola Plaza, MS6105,
Los Angeles, CA 90095\\
&\\
% --- Section: Education ---

\mycolor{Employment} 
& \textbf{University of California, Los Angeles} \hfill Los Angeles, CA \\ 
&\hspace{\myl} Hendrick Assistant Adjunct Professor\hfill  July 2025 --- June 2028  \\\vspace{0cm}\\
% --- Section: Education ---

\mycolor{Education} 
& \textbf{University of Minnesota, Twin Cities} \hfill Minneapolis, MN \\ 
&\hspace{\myl} Ph.D. in Mathematics\hfill  Sep 2020 --- July 2025  \\\vspace{0cm}
&\hspace{\myl} M.S. in Mathematics\hfill  Sep 2020 --- Jan 2024  \\\vspace{0cm}
%&\hspace{\myl} \hspace{\myl} \\ \vspace{0cm}
&\hspace{\myl} B.S. in Mathematics\hfill Sep 2016 --- May 2020\\
&\hspace{\myl} B.A. in Quantitative Economics\hfill Sep 2016 --- May 2020
\\
&\\

%{\mycolor{Awards and}} 
%& UROP Scholarship for undergrad research (UMN) \hfill 2020\\
%{\mycolor{scholarships}} 
%& Ordway Fellowship (School of Mathematics, UMN) \hfill 2024 \\
%& Doctoral Dissertation Fellowship (UMN) \hfill 2024 \\
%& \\



% --- Section: Research interests ---

\nohyphens{\mycolor{Research}}
& My research is in the area of algebraic combinatorics, often among the intersection of total positivity, geometry of the flag variety, and representation theory of linear groups. In particular, I am interested in (i) Schubert calculus and symmetric functions (ii) combinatorial representation theory and (iii) cluster algebras. \\
& \\



\nohyphens{\mycolor{Publications}} 
&\paper{1}{An Expansion Formula for Decorated Super-Teichm\"uler Spaces.} \\
&\ind  with  G. Musiker \& N. Ovenhouse. \\
&\ind  \href{https://doi.org/10.3842/SIGMA.2021.080}{SIGMA} {17} (2021) 080. \href{https://arxiv.org/abs/2102.09143}{arXiv 2102.09143}\\

&\paper{2}{Arborescence of Covering Graphs.} \\
&\ind  with S. Chepuri, C. Dowd, A. Hardt, G. Michel, \& V. Zhang.\\
&\ind     \href{https://doi.org/10.5802/alco.212}{Algebr. Comb.} Vol. 5 (2022) \href{https://arxiv.org/abs/1912.01060}{arXiv 1912.01060}\\

&\paper{3}{Double Dimer Covers on Snake Graphs from Super Cluster Expansions.}\\
&\ind  with G. Musiker \& N. Ovenhouse.\\
&\ind  \href{https://doi.org/10.1016/j.jalgebra.2022.05.033}{J. Algebra} Vol 608 (2022) pp. 325-381. \href{https://arxiv.org/abs/2107.14785}{arXiv 2107.14785}\\

&\paper{4}{Rowmotion Orbits of Trapezoid Posets.}\\
&\ind  with J. Wellman, Q. Dao, \& C. Yost-Wolff.\\
&\ind   \href{https://doi.org/10.37236/9769}{Electron. J. Comb.} 29-2 (2022) \href{https://arxiv.org/abs/2002.04810}{arXiv 2002.04810}\\

&\paper{5}{Rooted Clusters for Graph LP Algebras.}\\
&\ind  with E. Banaian, S. Chepuri, \& E. Kelley.\\
&\ind SIGMA 18 (2022), 089. \href{https://arxiv.org/abs/2107.14785}{arXiv 2107.14785}. \\

&\paper{6}{Matrix Formulae for Decorated Super Teichm\"uler Spaces.} \\
&\ind  with  G. Musiker \& N. Ovenhouse. \\
&\ind  J. Geom. Phys. (2023) \href{https://arxiv.org/abs/2208.13664}{arXiv 2208.13664}. \\

&\paper{7}{A Lattice Model for Super LLT Polynomials.} \\
&\ind  with M. Curran, C. Frechette, C. Yost-Wolff, \& V. Zhang\\
&\ind  Comb. Theory (2023) \href{https://arxiv.org/abs/2110.07597}{arXiv 2110.07597}. \\

&\\

\nohyphens{\mycolor{Preprints}}
&\paper{8}{Higher Dimer Covers on Snake Graphs.}\\
&\ind  with  G. Musiker, N. Ovenhouse, \& R. Schiffler. \\
&\ind \href{https://arxiv.org/abs/2306.14389}{arXiv 2306.14389}\\

&\paper{9}{Snake Graphs for Graph LP Algebras.}\\
&\ind  with E. Banaian, S. Chepuri, \& E. Kelley.\\
&\ind\href{https://arxiv.org/abs/2312.12313}{	arXiv:2312.12313}. \\

&\paper{10}{Higher $q$-Continued Fractions}\\
&\ind  with A. Burcroff, N. Ovenhouse, \& R. Schiffler.\\
&\ind\href{https://arxiv.org/abs/2408.06902}{	arXiv:2408.06902}. \\

&\paper{11}{Schubert Calculus and the Heisenberg Algebra}\\
&\ind\href{https://arxiv.org/abs/2409.20389}{	arXiv:2409.20389}. \\

&\\
\nohyphens{\mycolor{Proceedings}}
&\paper{12}{Rooted Clusters for Graph LP Algebras}\\
&\ind  with E. Banaian, S. Chepuri \& E. Kelley\\
&\ind \href{https://www.mat.univie.ac.at/~slc/wpapers/FPSAC2022/52.pdf}{proceeding of FPSAC 2022}\\
&\paper{13}{Double Dimers and Super Ptolemy Relations}\\
&\ind  with G. Musiker \& N. Ovenhouse\\
&\ind \href{https://www.mat.univie.ac.at/~slc/wpapers/FPSAC2023/79.pdf}{proceeding of FPSAC 2023}\\
%\nohyphens{\mycolor{In Preparation}}
%&\paper{11}{Scannable Divides of Finite Mutation Type}\\
%&\ind  with P. Dangwal, R. Lych, S. Nguyen, \& E. Pesikoff.\\
%%&\paper{9}{Greene-Kleitman Correspondence for Affine Posets.}\\
%%&\ind   with P. Pylyavskyy.\\
%&\paper{12}{Snake Graphs and Positivity for Graph LP Algebras.}\\
%&\ind   with E. Banaian, S. Chepuri, \& E. Kelley.\\

%&\paper{13}{Two Dimensional Fermions and Schubert Calculus}\\




&\\
% --- Section: Talks ---

{\mycolor{Invited Talks}} 
& \textbf{Schubert Calculus and Boson-Fermion Correspondence}  \\
&\ind  Combinatorics Seminar, UMN. \hfill February 2024\\
&\ind  Algebra Geometry and Combinatorics Seminar, UIUC. \hfill February 2024\\
&\ind  Combinatorics Seminar, UCSD. \hfill Mar 2024\\

& \textbf{The Greene-Kleitman Correspondence} \hfill November 2022 \\
&\ind  Student Algebra and Representation Seminar, SUNY Rutgers. \\
& \textbf{Super Cluster Algebras from Surfaces.}\\
&\ind  Combinatorics Seminar, University of Minnesota.  \hfill September 2022 \\
&\ind  AMS sectional meetings, Cincinnati.  \hfill April 2023 \\
& \textbf{Combinatorial Formulas for Graph LP algebras.} \hfill April 2022 \\
&\ind  Student Combinatorics and Algebras Seminar, University of Minnesota\\
& \textbf{Cluster Structures from Decorated Super-Teichm\"uller Spaces.}\hfill April 2022\\
&\ind Workshop on supergeometry and bracket structures, Fields institute.\\

& \textbf{Schur and LLT Polynomials from Lattice Models.} \hfill March 2021 \\
&\ind  Graduate Online Combinatorics Colloquium (GOCC) \\
& \textbf{$T$-paths Formula for Decorated Super-Teichm\"uller Spaces.} \hfill Feb 2021 \\
&\ind  Combinatorics Seminar, University of Minnesota \\

%& \\
& \\

%{\mycolor{Activities}} 
%& \textbf{Workshop on Supergeometry and Supermoduli} \hfill April 2023 \\
%&\ind  Simons Center for Geometry and Physics. \\
%& \textbf{Workshop on Symmetric Spaces.} \hfill August 2022 \\
%&\ind  University of Ottawa. \\
%& \textbf{The LA Workshop on Representation Theory and Geometry.} \hfill June 2022 \\
%&\ind  University of Southern California. \\
%& \textbf{MN Research Workshop in Algebraic Combinatorics.} \hfill May 2022 \\
%&\ind  Co-organized at University of Minnesota. \\
%&\textbf{Open Problems in Algebraic Combinatorics.}\hfill May 2022\\
%&\ind University of Minnesota.\\
%&\textbf{Workshop on Supergeometry and Bracket Structures.} \hfill March 2022\\
%&\ind Fields Institute, University of Toronto.\\
%%& \\
%& \\

% --- Section: Teaching experience ---

{\mycolor{Teaching}} 
& \textbf{Instructor, UCLA}  \\
&\ind  Math 206a (Topics in Combinatorics): Fall 2025\\

& \textbf{Teaching assistant, University of Minnesota}  \\
&\ind  Math 2263 (multivariable calculus): Spring 2021 Fall 2022\\
&\ind  Math 1372 (calculus 2): Fall 2021\\
&\ind  Math 1271 (calculus 1): Fall 2020 Spring 2021\\
&\ind  Math 1051 (pre-calculus): Fall 2019 Spring 2020\\
& \textbf{Instructor, University of Minnesota}  \\
&\ind  Math 4242 (applied linear algebra): Summer 2024\\
\\
% --- Section: Mentoring experience ---
{\mycolor{Mentoring}} 
&\textbf{PKU Algebraic Combinatorics Experience.}\\
&\ind $\bullet$ Unimodality of higher $q$-continued fractions (Mentor) \hfill Summer 2023\\
&\ind $\bullet$ Two conjectures on covering graphs (Mentor) \hfill Summer 2023\\
&\ind $\bullet$ Combinatorics of $\text{SL}_3$ cluster variables. (Mentor) \hfill Summer 2023\\
&\ind $\bullet$ Promotion orbit on trapezoid posets. (Mentor) \hfill Summer 2023\\
& \textbf{Algebra and Combinatorics REU at the University of Minnesota}  \\
&\ind $\bullet$ Classification for Divides of Finite Mutation Type. (TA)  \hfill Summer 2022\\
&\ind $\bullet$ Minimal Matching for $\text{dP}_3$ cluster algebras. (TA) \hfill Summer 2022\\
&\ind  $\bullet$ Kazhdan-Lusztig immanants and $\%$-immanants. (TA) \hfill Summer 2021\\
& \\

{\mycolor{Organization}} 
& \textbf{Minnesota Research Workshop in Algebra \& Combinatorics}  \\
&\ind Co-organizer. \hfill{Summer 2022 and 2023} \\
& \textbf{Student Combinatorics and Algebra Seminar at UMN}  \\
&\ind Co-organizer. \hfill {Fall 2023 -- Spring 2024}\\
& \textbf{Peking Algebraic Combinatorics Experience (REU).}\\
&\ind Co-organizer and problem mentor.\hfill{Summer 2023 and 2024}
\\
% --- Section: Various skills (programming, software, languages, etc.) ---

%{\mycolor{Skills}} 
%& \textbf{Programming}\\
%&\ind  Python, Mathematica, SageMath, HTML, \LaTeX. (fluent) \\
%&\ind C++, JavaScript, Julia, R. (intermediate)\\
%& \textbf{Languages} \\
%&\ind  Chinese Mandarin (native), English (fluent) \\
%& \\


% --- End of CV! ---
\end{longtable}

\end{document}
